Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-11-08 Ben_Manuscripts/structure_paper/Outline.tex
Equilibration scheme for bcc phase system. Should work on bridges
New figures and updates to outline

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 has been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in a low flux membrane, slowing research in the field.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated
  \end{itemize}

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling has been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

@@

   (4) Is it necessary to include any water in order to appropriately model the 
   Col\testsubscript{h} phase?
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a dry and "wet" system
   \end{itemize}

@@

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system. 
  \begin{itemize}
    \item Previous work has shown a coarse grain model self assemble into
    the H\textsubscript{II} phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.
    \item We attempted atomistic self-assembly by packing monomers into a box 
    using Packmol \cite{martinez_packmol:_2009}.
    \item Simulations of greater than 100 ns show no indicators of progress 
    towards an ordered system.  %BJC: include more details in supplment?
    \item To bypass the slow self-assembly process, python scripts are used
    to assemble monomers into a structure close to one of a number of 
    hypothesized equilibrium configurations (Figure~\ref{fig:python}).
  \end{itemize}

@@

  We used experimental wide angle X-ray scattering (WAXS) data (produced as described 
  in~\cite{feng_scalable_2014}) and small angle X-ray scattering (SAXS) data from~\cite{feng_thin_2016}
  to inform our choices of some initial structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily
  on the 2D WAXS data since it encodes all structural details down to the sub-nm scale.
  \begin{itemize}
        \item There are five major features of interest present in the 2D experimental
        pattern shown in Figure~\ref{fig:WAXS}.
        \item The first is located at $q_z$ = 1.7 \angstrom$^-1$,
        corresponding to a real spacing of 3.7 \angstrom~. The reflection is
        attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
        perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
        reflection will be referred to as R-$\pi$.
        \item A weak intensity line is located at exactly half the $q_z$ value of
        R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a
        real space periodic spacing of 7.4 \angstrom~. This reflection has been
        interpreted  as 2\textsubscript{1} helical ordering of aromatic rings
        along the z axis meaning if the positions of the aromatic rings can
        be traced by a helix, then for each turn in the helix, there should be
        two aromatic rings. For this reason it will be referred to as R-helix.
        \item A third major reflection is marked by a low intensity ring located
        at r = 1.4 \angstrom$^-1$. The real space separation
        corresponds to 4.5 \angstrom~ which is characteristic of the average
        spacing between packed alkane chains. This reflection will be called R-alkanes.
        \item Within R-alkanes, are four spots of higher relative intensity which
        will be called R-spots. All are located $\approx 40$ degrees from the $q_z$ axis
        in their respective quadrants. In many liquid crystal systems this can be
        explained by the tilt angle of the alkane chains with respect to the xy plane. % BJC: References
        \item The first corresponds to the spacing and symmetry of
        the d\textsubscript{100} plane which can be related to the distance between
        pores. The feature, which will be called R-pores, is characterized by dots
        along $q_z$ = 0. The spacing between dots is indicative of the hexagonal
        symmetry of the packed pores. The same information at higher resolution is obtained using a SAXS
        setup. By radially integrating the 2D data one gets a 1D curve which is 
        shown in Figure~\ref{fig:SAXS}.  % BJC: I also have 2D SAXS data
  \end{itemize}

  \begin{figure}[!ht]
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
        %       \vspace{12mm}
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \raisebox{.2\textwidth}{%
                \includegraphics[width=\linewidth]{WAXS_soft_confined.png}
                }
                \caption{}\label{fig:WAXS}
        \end{subfigure}
        \caption{(a) 2D wide angle X-ray scattering gives details about repeating
        features less than 1 nanometer apart. (b) 1D small angle X-ray scattering
        indicates hexagonal packing of pores as well as the spacing between pores.}\label{fig:SWAXS}
  \end{figure}

@@

%  We based the pore radii of our initial configuration on past TEM and size exclusion 
%  rejection data
  We used experimental Transmission electron microscopy (TEM) and size exclusion rejection
  data \cite{feng_scalable_2014,feng_thin_2016,zhou_supported_2005} to inform our definition
  of pore radius in the initial configuration.
  \begin{itemize}
    \item Experimental evidence suggests uniform pores with radii of 0.6 nm 
    \item Comparing a geometric measurement of pore size derived from an atomistic model,
    to a less precise, experimentally derived pore size estimate, will give ambiguous results.
    \item What is meant by pore radius will not be clear until we establish a clear picture
    of the nanoscopic pore environment.
    \item When constructing pores, we chose the carboxylate carbon from the monomer
    head group as a reference atom, and placed it a distance r from the pore center,
    where r is the pore radius. (See Supplemental Information) %BJC: figure making that clear
    \item We will not make direct comparisons of pore radius between our model 
    and experiment to avoid the ambiguity, however, we do define a pore radius  
    for our own purposes.
    %\item To measure the pore radius in our model we calculate the distance between 
    %the center of mass of each aromatic ring and the center of mass of all aromatic rings
    %in their respective pores.
  \end{itemize}

@@

  %We tested a number of equilibration schemes of increasing complexity in an effort
  %to match experiment while overcoming kinetic limitations that might trap the system 
  %in a metastable states. All start from an initial configuration generated with 
  %chosen structure variables.
  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          slowly released until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate further in the NPT
          ensemble for at least 400 ns.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          %an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All waters placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental?
  \end{itemize}

@@

  We calculated ionic conductivity using two different methods for robustness.
  \begin{itemize}
    \item The Nernst-Einstein relation relates the DC ionic conductivity to 
    ion diffusivity, $D$, concentration, $C$, ion charge, $q$, the boltzmann 
    constant, $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$  %BJC: need citation
    \item Sodium ion diffusion coefficients were found by calculating the slope
    of the linear region of the mean square displacement curve as indicated by
    the einstein relation \cite{einstein_investigations_1956}.]
    %MRS: state how you decided which was the linear region.
    %BJC: nothing fancy other than looking at the msd plot and determining where to
    % start fitting. I suppose I could minimize R^2 but that seems like overkill
    \item We looked at the MSD plot to determine where to begin and end a linear fit
    \item Ion concentration was measured with respect to the entire unit cell. 
    \item The second method, termed the 'Collective Diffusion' model, measures 
    the movement of the collective variable, Q, which is defined as the amount
    of charge transfer through the system and can be thought to represent
    the center of charge of the system.
    \item The conductance, $\gamma$ of the system can be calculated as:
    $$ \gamma = \dfrac{D_Q}{k_b T} $$ Conversion to ionic conductivity is
    achieved by multiplying by channel length and dividing by the membrane
    cross sectional area.
    \item $D_Q$ is the diffusion coefficient of the collective variable Q. It can
    be calculated using the einstein relation.
    \item A full derivation of the model can be accessed elsewhere \cite{liu_collective_2013}.
  \end{itemize}

@@

  To discern the composition of the monomer layers, addressing (1), we ran
  simulations of systems created with 4 - 8 monomers per layer.  % BJC: can't figure out how to word this : should be 'with between 4 and 8 monomers' or 'with from 4 to 8 monomers'?
  \begin{itemize}
        \item Both the sandwiched and parallel displaced configurations were tested.
        \item All systems are stable after 400 ns of simulation.  % BJC: edited this for clarity
        %MRS: you only ran for 50 ns, or after 50 ns some of them started to destabilize?  Is the point you are making that they are ALL pretty stable, or all at least a little stable.   
        %BJC: Working on getting them all run out to 400 ns
        \item Table ~\ref{table:p2p} shows the pore spacing for all systems tested.
        \item Systems built with 5 monomers in each layer equilibrate to a pore spacing
        that is most consistent with the experimental value of 4.12 nm derived from
        SAXS measurements (Figure~\ref{fig:SAXS}).
        \item The remainder of this discussion will focus on the analysis of systems
        built with 5 monomers per layer.
  \end{itemize}

  To answer (2), we verified that the system stays partitioned into layers by
  plotting the pair correlation function, $g(z)$ calculated between aromatic rings along
  the length of the pores (Fig~\ref{fig:zdf}).
  \begin{itemize}
        \item Sandwiched configuration layers stack 4.29 \angstrom apart while
        parallel displaced configuration layers stack 4.37 \angstrom apart.
        The power spectrums used to calculate the reported layer spacing are given in the 
        Supplemental Information.  %BJC: "Power spectrum" and okay description here?
	\item Even though the layer spacing is similar, the sandwiched configuration
	exhibits more defined layers with its first peak deviating from the mean 
        number density by 61 \%. %BJC: I mention how I do this calculation in the methods. Is it clear that I'm doing that here?
	\item The parallel displaced configuration deviates from the mean by only 
	12 \% in comparison.
  \end{itemize}

@@

  \begin{table}[h]
  \centering
  \begin{tabular}{ccc}
  \toprule
  		   & \multicolumn{2}{c}{Starting Configuration} \\
  \hline
  Monomers per layer & Sandwiched & Parallel Displaced \\
  \midrule
  4 & $3.71 \pm 0.04$ & $3.84 \pm 0.02$ \\
  5 & $4.20 \pm 0.04$ & $4.23 \pm 0.04$ \\
  6 & $4.83 \pm 0.03$ & $4.85 \pm 0.02$ \\
  7 & $4.73 \pm 0.03$ & $4.84 \pm 0.03$ \\
  8 & $5.08 \pm 0.04$ & $5.46 \pm 0.03$ \\
  \bottomrule
  \end{tabular}
  \caption{The pore spacing (given in nm) of the model increases as
  number of monomers in each layer increases. The pore spacing of a system
  starting in the sandwiched configuration is systematically lower than 
  that started in an offset configuration. Systems built with 5 monomers
  per layer in a parallel displaced configuration result in a pore spacing
  closest to the experimental 4.12 nm}
  \label{table:p2p} 
  \end{table}

@@

  We answer question (3) by simulating X-ray diffraction patterns produced from
  equilibrated MD trajectories.
  %We created systems using three distinct initial configurations and examined
  %their structure by simulating X-ray diffraction patterns produced from equilibrated MD trajectories.
  \begin{itemize}
	\item We created configurations in both the parallel displaced and sandwiched
	configurations
        \item The systems were assembled and equilibrated according to our dry equilibration procedure
        \item Simulated diffration patterns were generated using portions of the
        trajectory after equilibration.
        \item We assume the simulation has equilibrated when the distance between
        pores and the membrane thickness stopped changing. %BJC: might be better placed in methods
        \item Simulated diffraction patterns for both structures are shown and 
	compared to experimenta in Figure ~\ref{fig:XRDsim}.
  \end{itemize}
%

@@

  We answer (4) and further support (3) by preparing parallel displaced and
  sandwiched configurations according to the wet equilibration procedure.
  \begin{itemize}
	\item We do not have an experimental measurement of water in the pores
	\item We decided the number of waters to put in each pore by estimating
        the number of hydrogen bonds in which pore entities could participate.
	\item In solution, each headgroup has a charge of -1
	\item We estimate there to be about 4 water molecules per layer based 
	on various packing configurations and geometric constraints
	\item 4 water molecules per layer translates to about 1.6 wt \%
  \end{itemize}

>>>>

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 have been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    % BJC: New paragraph?
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in low flux membranes, slowing research in the field.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated.
  \end{itemize}

  % BJC: Not sure this figure needs to be in the paper
  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{soft_confinement.png}
	\caption{The wedge shaped liquid crystal monomer (a) self assembles into 
	mesophases with hexagonally packed pores (b). The pores are made of 
	stacked monomer disks. A sub-micron-thick film is created by casting a 
	dilute solution of Na-GA3C11/THF solution onto a silicon substrate and 
	and allowing the solvent to evaporate. The thin film contains nanoporous 
	columns which lie parallel to the film plane. (d) When a soft PDMS pad is 
	imposed to the thin film, with subsequent thermal annealing, the columns
	align perpendicular to the film plane. (e) Photo-cross-linking of the
	aligned film creates a mechanically stable thin film with vertically 
	aligned nanopores}~\label{fig:soft}
  \end{figure}

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  \begin{figure}[ht]
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport}~\label{fig:detail}
  \end{figure}

@@

   (4) Is it necessary to include any water in order to appropriately model the 
   Col\testsubscript{h} phase?
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
   \end{itemize}

@@

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system. 
  \begin{itemize}
    \item Previous work has shown a coarse grain model self assemble into
    the H\textsubscript{II} phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.
    \item We attempted atomistic self-assembly by packing monomers into a box 
    using Packmol \cite{martinez_packmol:_2009}.
    \item Simulations of greater than 100 ns show no indicators of progress 
    towards an ordered system.  %BJC: include more details in supplment
    \item To bypass the slow self-assembly process, python scripts are used
    to assemble monomers into a structure close to one of a number of 
    hypothesized equilibrium configurations (Figure~\ref{fig:python}).
  \end{itemize}

@@

  We used experimental wide angle X-ray scattering (WAXS) data (produced as described 
  in~\cite{feng_scalable_2014}) and small angle X-ray scattering (SAXS) data from~\cite{feng_thin_2016}
  to inform our choices of some initial structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily
  on the 2D WAXS data since it encodes all structural details down to the sub-nm scale.
  \begin{itemize}
        \item There are five major features of interest present in the 2D experimental
        pattern shown in Figure~\ref{fig:WAXS}.
        \item The first is located at $q_z$ = 1.7 \angstrom$^-1$,
        corresponding to a real spacing of 3.7 \angstrom~. The reflection is
        attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
        perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
        reflection will be referred to as R-$\pi$.
        \item A weak intensity line is located at exactly half the $q_z$ value of
        R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a
        real space periodic spacing of 7.4 \angstrom~. This reflection has been
        interpreted  as 2\textsubscript{1} helical ordering of aromatic rings
        along the z axis meaning if the positions of the aromatic rings can
        be traced by a helix, then for each turn in the helix, there should be
        two aromatic rings. For this reason it will be referred to as R-helix.
        \item A third major reflection is marked by a low intensity ring located
        at r = 1.4 \angstrom$^-1$. The real space separation
        corresponds to 4.5 \angstrom~ which is characteristic of the average
        spacing between packed alkane chains. This reflection will be called R-alkanes.
        \item Within R-alkanes, are four spots of higher relative intensity which
        will be called R-spots. All are located $\approx 40$ degrees from the $q_z$ axis
        in their respective quadrants. In many liquid crystal systems this can be
        explained by the tilt angle of the alkane chains with respect to the xy plane. % BJC: References
        \item The final feature corresponds to the spacing and symmetry of
        the d\textsubscript{100} plane which can be related to the distance between
        pores. The feature, which will be called R-pores, is characterized by dots
        along $q_z$ = 0. The spacing between dots is indicative of the hexagonal
        symmetry of the packed pores. The same information at higher resolution is obtained using a SAXS
        setup. By radially integrating the 2D data one gets a 1D curve which is 
        shown in Figure~\ref{fig:SAXS}.  % BJC: I also have 2D SAXS data
  \end{itemize}

  \begin{figure}[!ht]
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
        %       \vspace{12mm}
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \raisebox{.2\textwidth}{%
                \includegraphics[width=\linewidth]{WAXS_annotated.png}
                }
                \caption{}\label{fig:WAXS}
        \end{subfigure}
        \caption{(a) 1D small angle X-ray scattering indicates hexagonal packing of
	pores as well as the spacing between pores (b) 2D wide angle X-ray scattering 
	gives details about repeating features less than 1 nanometer apart}\label{fig:SWAXS}
  \end{figure}

@@

%  We based the pore radii of our initial configuration on past TEM and size exclusion 
%  rejection data
  We used experimental Transmission electron microscopy (TEM) and size exclusion rejection
  data \cite{feng_scalable_2014,feng_thin_2016,zhou_supported_2005} to inform our definition
  of pore radius in the initial configuration.
  \begin{itemize}
    \item Experimental evidence suggests uniform pores with radii of 0.6 nm 
    \item Comparing a geometric measurement of pore size derived from an atomistic model,
    to a less precise, experimentally derived pore size estimate, will give ambiguous results.
    \item What is meant by pore radius will not be clear until we establish a clear picture
    of the nanoscopic pore environment.
    \item When constructing pores, we chose the carboxylate carbon from the monomer
    head group as a reference atom, and placed it a distance r from the pore center,
    where r is the pore radius (~\ref{fig:pore_radius_illustration}). % BJC: Figure in Supplemental Information?
    \item We will not make direct comparisons of pore radius between our model 
    and experiment to avoid the ambiguity, however, we do define a pore radius  
    for our own purposes.
    %\item To measure the pore radius in our model we calculate the distance between 
    %the center of mass of each aromatic ring and the center of mass of all aromatic rings
    %in their respective pores.
  \end{itemize}

  \begin{figure}[ht]
 	\centering
 	\includegraphics[width=\textwidth]{pore_radius_illustration.png}
	\caption{The pore radius of the initial configuration is defined using the 
        carboxylate carbon as a reference}~\label{fig:pore_radius_illustration}
  \end{figure}

@@

  %We tested a number of equilibration schemes of increasing complexity in an effort
  %to match experiment while overcoming kinetic limitations that might trap the system 
  %in a metastable states. All start from an initial configuration generated with 
  %chosen structure variables.
  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          slowly released until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All waters placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental?
  \end{itemize}

@@

  We calculated ionic conductivity using two different methods for robustness.
  \begin{itemize}
    \item The Nernst-Einstein relation relates the DC ionic conductivity to 
    ion diffusivity, $D$, concentration, $C$, ion charge, $q$, the boltzmann 
    constant, $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$  %BJC: need citation
    \item Sodium ion diffusion coefficients were found by calculating the slope
    of the linear region of the mean square displacement curve as indicated by
    the einstein relation \cite{einstein_investigations_1956}.]
    %MRS: state how you decided which was the linear region.
    %BJC: nothing fancy other than looking at the msd plot and determining where to
    % start fitting. I suppose I could minimize R^2 but that seems like overkill
    \item We looked at the MSD plot to determine where to begin and end a linear fit
    \item Ion concentration was measured with respect to the entire unit cell. 
    % new paragraph
    \item The second method, termed the 'Collective Diffusion' model, measures 
    the movement of the collective variable, Q, which is defined as the amount
    of charge transfer through the system and can be thought to represent
    the center of charge of the system.
    \item The conductance, $\gamma$ of the system can be calculated as:
    $$ \gamma = \dfrac{D_Q}{k_b T} $$ Conversion to ionic conductivity is
    achieved by multiplying by channel length and dividing by the membrane
    cross sectional area.
    \item $D_Q$ is the diffusion coefficient of the collective variable Q. It can
    be calculated using the einstein relation.
    \item A full derivation of the model can be accessed elsewhere \cite{liu_collective_2013}.
  \end{itemize}

@@

  To discern the composition of the monomer layers, addressing (1), we ran
  simulations of systems created with 4 - 8 monomers per layer.  % BJC: can't figure out how to word this : should be 'with between 4 and 8 monomers' or 'with from 4 to 8 monomers'?
  \begin{itemize}
        \item We created systems in both the parallel displaced and sandwiched configurations
	\item All systems were prepared according to the dry equilibration procedure
        \item All systems are stable after 400 ns of simulation.  
        \item Table ~\ref{table:p2p} shows the pore spacing for all systems tested.
        \item Systems built with 5 monomers in each layer equilibrate to a pore spacing
        that is most consistent with the experimental value of 4.12 nm derived from
        SAXS measurements (Figure~\ref{fig:SAXS}).
        \item The remainder of this discussion will focus on the analysis of systems
        built with 5 monomers per layer.
  \end{itemize}

  %BJC: maybe split into 2 paragraphs
  To answer (2), we verified that the system stays partitioned into layers by
  plotting the pair correlation function, $g(z)$ calculated between atoms along
  the length of the pores (Fig~\ref{fig:zdf}).
  \begin{itemize}
	\item We measured $g(z)$ with respect to aromatic rings in the head groups
	and, separately, with respect to carbon atoms on the alkane chains.
	\item In both cases, we see a degree of layer partitioning
        \item Using the head groups as reference, we see that sandwiched configuration
	layers stack 4.29 \angstrom apart while parallel displaced configuration layers
	stack 4.37 \angstrom apart.
	\item Using the alkane chain carbons as reference we see a layer spacing of 4.39 \angstrom
	and 4.38 \angstrom in the sandwiched and parallel displaced configurations respectively.
        \item The power spectrums used to calculate the reported layer spacings are given in the 
        Supplemental Information.  %BJC: "Power spectrum" an okay description here?
	\item We see a difference in the spacing in the head group versus tail region
	because there is more volume to fill with less densely packed carbon chains in the tail region. 
	% BJC: vvv Need to get the same information for the tail zdfs.
	\item Even though the layer spacing is similar, the sandwiched configuration
	exhibits more defined layers with its first peak deviating from the mean 
        number density by 61 \%. %BJC: I mention how I do this calculation in the methods. Is it clear that I'm doing that here?
	\item The parallel displaced configuration deviates from the mean by only 
	12 \% in comparison.
  \end{itemize}

@@

  \begin{table}[h]
  \centering
  \begin{tabular}{ccc}
  \toprule
  		   & \multicolumn{2}{c}{Starting Configuration} \\
  \hline
  Monomers per layer & Sandwiched & Parallel Displaced \\
  \midrule
  4 & $3.72 \pm 0.04$ & $3.84 \pm 0.02$ \\  % done
  5 & $4.20 \pm 0.04$ & $4.23 \pm 0.04$ \\
  6 & $4.69 \pm 0.04$ & $4.72 \pm 0.03$ \\
  7 & $4.66 \pm 0.02$ & $4.82 \pm 0.03$ \\  % done
  8 & $5.00 \pm 0.04$ & $5.33 \pm 0.02$ \\
  \bottomrule
  \end{tabular}
  \caption{The pore spacing (given in nm) of the model increases as
  number of monomers in each layer increases. The pore spacing of a system
  starting in the sandwiched configuration is systematically lower than 
  that started in an offset configuration. Systems built with 5 monomers
  per layer in a parallel displaced configuration result in a pore spacing
  closest to the experimental 4.12 nm}
  \label{table:p2p} 
  \end{table}

@@

  We answer question (3) by simulating X-ray diffraction patterns produced from
  equilibrated MD trajectories.
  %We created systems using three distinct initial configurations and examined
  %their structure by simulating X-ray diffraction patterns produced from equilibrated MD trajectories.
  \begin{itemize}
	\item Systems of 5 monomers per layer in the parallel displaced and sandwiched
	configurations were tested.
        \item Simulated patterns were generated using portions of simulation
        trajectory after equilibration.
        \item The patterns for both structures are shown and compared to experiment
	in Figure ~\ref{fig:XRDsim}.
  \end{itemize}
%

@@

  We answer (4) and further support (3) by preparing parallel displaced and
  sandwiched configurations according to the wet equilibration procedure.
  \begin{itemize}
	\item We do not have an experimental measurement of water in the pores
	\item We tested a range of water concentrations from 0.5 to 5 percent % working on a 0.5 wt % sytstem now just to see
	\item Our lower bound represents a system with on average 2 water 
	molecules for each monomer layer.
	\item Figure ~\ref{fig:solvation} shows results for each configuration.
%	\item Initially, we decided the number of waters to put in each pore by
%	estimating the number of hydrogen bonds in which pore entities could participate.
%	\item In solution, each headgroup has a charge of -1
%	\item We estimate there to be at least 2 water molecules per layer based 
%	on various packing configurations and geometric constraints
%	\item 4 water molecules per layer translates to about 1.6 wt \%
  \end{itemize}

  \begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{solvation.png}  % still working on formatting so this is a screenshot of a doctored latex figure
	\label{fig:solvation}
  \end{figure}

  In all cases, water disrupts structuring of the model.
  \begin{itemize}
	\item The plots in Figure ~\ref{fig:solvation} are normalized 
	to have the same colorbar as the dry systems. 
	\item The intensity of the reflections decrease when water is 
	added to the system.
	\item In systems built with 5 weight percent water, R-$\pi$ and
	R-spots become nearly indistinguishable from R-alkanes.
  \end{itemize}
